Protected Areas Practical

### 2024-11-25 tutorial/MSc_PA_practical_answers.Rmd
separated exercise and answers

# Exercise 4 (Bonus)

As a final exercise, you will "lock out" some areas from protection. These are places like cities that should be ignored from the prioritisation.

@@

```{r, fig.width=5, fig.height=7, echo=FALSE}
locked_out = terra::rast("data/locked_out.tif")

@@

@@

# check percentage of the features that have their target met given the solution
print(mean(p4_target_coverage$met) * 100)

@@

Notice that it reality it doesn't change much because we are using the same data to derive the cost layer and the

>>>>

# Exercise 4: Lock out land from conservation

As an exercise, you will "lock out" some areas from protection. These are places like cities that should be ignored from the prioritisation.

@@

```{r, fig.width=5, fig.height=7}
locked_out = terra::rast("data/locked_out.tif")

@@

Notice that it reality it doesn't change much because we are using the same data to derive the cost layer and the locked out layer. However locking areas out of the prioritisation means that some "expensive" but biodiverse sites cannot be protected, and need to be replaced with more and cheaper planning units, meaning that more land has to be protected in compensation...

```{r}

@@

# check percentage of the features that have their target met given the solution
print(mean(p4_target_coverage$met) * 100)
```

# Exercise 5: Implementing Marxan

Marxan addresses conservation problems the opposite way round to prioritizr. It tries to find the cheapest solution (that will impact the least people) while meeting the biodiversity target (here to protect 30% of the range of each species). This can be done by replacing `add_min_shortfall_objective(budget)` with `add_min_set_objective()` in the `p1` solution. We no longer rely on a budget.

```{r}
py <-
  problem(cost, features = species) %>%
  # add_min_shortfall_objective(budget) %>% # replace this
  add_min_set_objective() %>% # with this
  add_relative_targets(0.3) %>%
  add_binary_decisions() %>%
  add_locked_in_constraints(PAs) %>%
  add_default_solver(gap = 0.1, verbose = FALSE)

sy <- solve(py)

# changed the plot colours because why not
# plot the solution
par(mfrow=c(1,2))# split into 2 plots side by side
plot(s1, main = "maximise biodiversity", axes = FALSE)
plot(sy, main = "minimise cost", axes = FALSE)
par(mfrow=c(1,1))# go back to only 1 plot

@@

### 2024-11-25 tutorial/MSc_PA_practical_exercise.Rmd
separated exercise and answers

```{r}
# calculate budget
budget <- terra::global(cost, "sum", na.rm = TRUE)[[1]] * 0.3

# create problem
pt <-
  problem(cost, features = species) %>%
  add_min_shortfall_objective(budget) %>%
  add_relative_targets(0.7) %>%
  add_binary_decisions() %>%
  add_default_solver(gap = 0.1, verbose = FALSE)

@@

```{r}
pz <-
  p2 %>%
  add_boundary_penalties(penalty = 0.7, edge_factor = 0.2)

@@

```{r, fig.width=5, fig.height=7}
locked_out = terra::rast("data/locked_out.tif")

@@

# create new problem with boundary penalties added to it
p4 <-
  p2 %>%
  add_locked_out_constraints(locked_out)

@@

# plot the solution
par(mfrow=c(2,2))# split into 3 plots side by side
plot(s1, main = "Baseline", axes = FALSE)
plot(s2, main = "With PAs", axes = FALSE)
plot(s4, main = "With PAs + Avoid cities", axes = FALSE)
plot(s3, main = "With PAs + Boundary layer", axes = FALSE)
par(mfrow=c(1,1))# go back to only 1 plot

@@

```{r}

# calculate target coverage for the solution
p4_target_coverage <- eval_target_coverage_summary(p4, s4)
print(p4_target_coverage)

@@

```{r}
py <-
  problem(cost, features = species) %>%
  # add_min_shortfall_objective(budget) %>% # replace this
  add_min_set_objective() %>% # with this
  add_relative_targets(0.3) %>%
  add_binary_decisions() %>%
  add_locked_in_constraints(PAs) %>%
  add_default_solver(gap = 0.1, verbose = FALSE)

@@

# changed the plot colours because why not
# plot the solution
par(mfrow=c(1,2))# split into 2 plots side by side
plot(s1, main = "maximise biodiversity", axes = FALSE)
plot(sy, main = "minimise cost", axes = FALSE)
par(mfrow=c(1,1))# go back to only 1 plot

>>>>

```
# calculate budget
budget <- terra::global(cost, "sum", na.rm = TRUE)[[1]] * 0.3

# create problem
pt <-
  problem(cost, features = species) %>%
  add_min_shortfall_objective(budget) %>%
  add_relative_targets(0.3) %>% # change this line
  add_binary_decisions() %>%
  add_default_solver(gap = 0.1, verbose = FALSE)

@@

```
pz <-
  p2 %>%
  add_boundary_penalties(penalty = 0.003, edge_factor = 0.5) # change the values here

@@

```
locked_out = terra::rast("data/locked_out.tif")

@@

# create new problem with boundary penalties added to it
p4 <-
  p2 %>%
  # add add_locked_out_constraints(locked_out) here

@@

# plot the solution
par(mfrow=c(2,2))# split into 3 plots side by side
plot(s1, main = "Baseline", axes = FALSE)
plot(s2, main = "With PAs", axes = FALSE)
plot(s4, main = "With PAs + Avoid cities", axes = FALSE)
plot(s3, main = "With PAs + Boundary layer", axes = FALSE)
par(mfrow=c(1,1))# go back to only 1 plot
```

@@

```
# calculate target coverage for the solution
p4_target_coverage <- eval_target_coverage_summary(p4, s4)
print(p4_target_coverage)

@@

```
py <-
  problem(cost, features = species) %>%
  add_min_shortfall_objective(budget) %>% # replace this with add_min_set_objective()
  add_relative_targets(0.3) %>%
  add_binary_decisions() %>%
  add_locked_in_constraints(PAs) %>%
  add_default_solver(gap = 0.1, verbose = FALSE)

@@

# changed the plot colours because why not
# plot the solution
par(mfrow=c(1,2))# split into 2 plots side by side
plot(s1, main = "maximise biodiversity", axes = FALSE)
plot(sy, main = "minimise cost", axes = FALSE)
par(mfrow=c(1,1))# go back to only 1 plot
```
